---
title: "Gupta_Aarushi_Lesson04"
author: "SID 905-328-941"
date: "Stats 20 Summer 20"
output: 
  html_document:
    toc: true
    toc_depth: 3
    theme: cerulean
    highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## 4 

### 1.(a)

```{r}
a <- matrix(c(18.1, 15.5, 14.7, 17.4, 12.6, 15.5, 15.2, 13.2), nrow = 2, ncol = 4, byrow = TRUE)
a
```

### 1.(b)

```{r}
## overwrite 2nd column with square root of original 2nd column
a[,2] <- sqrt(a[,2])
a
```

### 1.(c)

```{r}
## remove 1st row and 4th column and store as a new object
a2 <- a[2,1:3]
a2
is.vector(a2)
```
R returns a vector when we remove both the 1st row and 4th column from (b) and store it as a new object.

### 2.(a)

```{r}
load("~/Desktop/2020 Classes/STATS 20/Lesson 3/Lesson3SU.RData")
num2 <- list(Q4B, Q4C, Q4D331, Q4D430, Q4D531, Q4D614)
```

### 2.(b)

```{r}
## shows single list object using all data in Lesson3SU.RData
str(num2)
```

### 3

```{r}
MyModel <- lm(dist ~ speed, data = cars)
MySummary <- summary(MyModel)
mean(MySummary$residuals)
```


### 4

plot(MyModel) generates 4 different scatterplots, all involving residuals and one other variable. plot(MySummary) produces an error, even though it's a list. MySummary generates an error because there are no clearly defined x and y coordinates in the list for R to take the data from.