---
title: "Lesson 7 - Data Frames III"
author: "Aarushi Gupta SID 905-328-941"
date: "Stats 20 Section 1"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r}
library(dplyr)
library(readr)
```


## 2A

Using the COlevel and COmostrecent datasets, we can determine the correlation between the Covid-19 infection rate for each county and its population density (similar to what was in the Lesson 7 instructions). Specifically, we can figure this out by calculating the correlation coefficient, after combining the two datasets. 


## 2B

```{r 2B}
library(DT)
COlevel <- read_csv("COlevel.csv")
COmostrecent <- read_csv("COmostrecent.csv")
combined <- inner_join(COmostrecent, COlevel, by = c("state","county"))
datatable(combined)
combined <- combined %>% mutate(coInfrate = round(cases/(coPOP/10000), digits = 2))
datatable(combined)
```

Thus, we calculated each county's Covid-19 infection rate per 10,000 people (based on given population density).

## 3A

The COlevel dataframe is about each county's population, land area, and demographic breakdown, and was collected prior to the COmostrecent dataframe. The COmostrecent dataframe only measures coronavirus-related county data. COlevel was probably part of some US census and is thus required to have detailed information on every single county, whereas COmostrecent is newer and based solely on county reporting. In short, COlevel includes population and demographic data which is required information for all counties, but Covid-19 data is not. 

## 3B

```{r 3B}
library(DT)
STfeb20 <- read_csv("STfeb20.csv")
STmar20 <- read_csv("STmar20.csv")
dt1 <- left_join(STfeb20, STmar20, by = "state")
datatable(dt1)
dt2 <- right_join(STfeb20, STmar20, by = "state")
datatable(dt2)
```

The number of rows increased dramatically as we moved from January to February to March 2020, as shown by the data increasing from 4 to 10 to 55 rows. This makes sense because Covid-19 cases also increased in most states during this time period. 

## 3C

```{r 3C}
dt3 <- left_join(COlevel, COmostrecent, by = c("state","county"))
dim(dt3)
dt4 <- right_join(COlevel, COmostrecent, by = c("state","county"))
dim(dt4)
```

left_join results in 3143 observations, whereas right_join results in 3060 observations, because left_join is dependent on the number of rows in COlevel and right_join is dependent on the number of rows in COmostrecent.
left_join provides more data, since it displays the population, area, and demographic data of each county, and also the Covid-19 data if reported. right_join only reports the population, area, and demographic data of counties who have reported Covid-19 data as well. 
