---
title: "Lesson 11 - Tables"
author: "Aarushi Gupta SID 905-328-941"
date: "Stats 20 Section 1"
output: html_document
---

```{r}
load(url("http://www.stat.ucla.edu/~vlew/datasets/ucpay2017.RData"))
library(knitr)
library(dplyr)
library(pander)
```

# 2.  Typical tabling questions from a SQL interview

Just write "yes" or "no" next to each question and if you wrote "yes", please indicate whether you could program it in base R, dplyr or both:

### A.1 minimum "RegularPay" and 

yes, Base R and dplyr.

### A.2 maximum "RegularPay" 

yes, Base R and dplyr.

### B. Count all employees whose Position contains the string 'Prof'  (and not 'Profl'):

yes, Base R and dplyr.

### C.1 total "Total Wages", 

yes, dplyr.

### C.2 mean "Total Wages" 

yes, dplyr.

### C.3 median "Total Wages" 

yes, dplyr.

### D  Find the total and mean "RegularPay" for each unique Position which contains the string 'Prof' (and not 'Profl') and also has at least (at minimum) 40 employees (40 rows, observations).

yes, dplyr.

# 3. Optional

### A.

yes, dplyr.

### B.

yes, dplyr.

### C.

yes, dplyr.

# 5. See how many you can answer now that you've read the entire lesson

### A.1 minimum "RegularPay" and 

```{r}
# Base R
min(uclaPayroll2017$RegularPay)

# dplyr
uclaPayroll2017 %>% summarize(minimum = min(RegularPay))
```

### A.2 maximum "RegularPay" 

```{r}
# Base R
max(uclaPayroll2017$RegularPay)

# dplyr
uclaPayroll2017 %>% summarize(maximum = max(RegularPay))
```

### B. Count all employees whose Position contains the string 'Prof'  (and not 'Profl'):

```{r}
uclaPayroll2017 %>% filter(substr(`Position`, 1, 4) == "Prof") %>% 
    summarize(count = n())

```

### C.1 total "Total Wages", 

```{r}
uclaPayroll2017 %>% filter(Position == "Prof") %>% 
  summarise(total = sum(`Total Wages`)) %>% 
  pander()
```

### C.2 mean "Total Wages" 

```{r}
uclaPayroll2017 %>% group_by(`Department Title`) %>% 
  summarise(total = sum(`Total Wages`), 
            avg = mean(`Other Pay (Payroll Explorer)`),
            minimum = min(`Other Pay (Payroll Explorer)`), 
            maximum = max(`Other Pay (Payroll Explorer)`)) %>% 
  pander()
```

### C.3 median "Total Wages" 

```{r}
uclaPayroll2017 %>% group_by(`Department Title`) %>% 
  summarise(total = sum(`Total Wages`), 
            avg = mean(`Other Pay (Payroll Explorer)`),
            minimum = min(`Other Pay (Payroll Explorer)`), 
            maximum = max(`Other Pay (Payroll Explorer)`)) %>% 
  pander()
```

### D  Find the total and mean "RegularPay" for each unique Position which contains the string 'Prof' (and not 'Profl') and also has at least (at minimum) 40 employees (40 rows, observations).

```{r}
load(url("http://www.stat.ucla.edu/~vlew/datasets/ucpay2017.RData"))
library(knitr)
library(dplyr)
library(pander)
```
